---
title: "CAP Task Events"
author: "Elena Peterson"
date: "5/25/2021"
output: html_document
---

Script for merging fmri task files (nback_raw_merge_[date].csv) and generating volume-by-volume list of task events (CAP_events_neut_[date].csv).

If task data files and events of importance haven't changed, no need to run this again.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Read in & merge all task files (*skip if this has already been done*):
```{r}
setwd("~/Desktop/Dissertation/Data/RRAY_scantask_data/")

alldata = list.files(recursive=T,pattern="*.csv")
alldata
alldata = lapply(alldata, read.csv, header=TRUE) #finicky error... just try again.
mydata = Reduce(function(...)merge(...,all=T),alldata)

# add in fixed data (tasks where subjects switched keys)
setwd("~/Desktop/Dissertation/Data/RRAY_scantask_data_fixed/")

#list.files(recursive=T,pattern="*.csv")
alldata2 = list.files(recursive=T,pattern="*.csv")
alldata2
alldata2 = lapply(alldata2, read.csv, header=TRUE) #finicky error... just try again.
mydata2 = Reduce(function(...)merge(...,all=T),alldata2)

# merge raw and fixed data:
mydata3 = merge(mydata, mydata2, all=T)

# save file
write.csv(mydata3, file = paste0("~/Desktop/Dissertation/Data/Brain_Data/nback_raw_merge_", Sys.Date(),".csv"),row.names=FALSE)

```

Read in data files:
```{r}
#use merged data file:
mydata <- read.csv("../Data/Brain_Data/nback_raw_merge_2024-03-29.csv",header=TRUE)

#subject list:
dsubs = read.csv("../Data/Brain_Data/vols_per_sub.csv",header=TRUE)
names(dsubs)

```

Next, create list of task events along with corresponding fmri volumes (*skip if this has already been done*):
```{r}
# tv = time to collect one volume
tv = 800 #(based on specs)

#Cumulative time per volume:
voltimes <- c(1:720)*tv
n = data.frame("voltimes"=voltimes)
e = data.frame("voltimes"=voltimes)

#Determine events at each volume
#use all data, turn into loop to get average times

gtimes =  data.frame("ind"=1:48) #48= number events

for (f in 1:length(alldata)) {
  #Get scanner pulse time:
  #SyncSlide.OnsetTime?
  d = alldata[[f]]
  pulsetime <- d$trial_startT[3] #check....
  
  #list of all block start times
  btimes <- d$trial_startT[which(d$trials.thisN==0)]
  
  #list of other starts
  cuetimes<- d$trial_startT[which(d$trials.thisN==0)-1]
  endtimes<- d$trial_startT[which(d$trials.thisN==9)+1]
  
  alltimes <- c(btimes,cuetimes,endtimes)
  alltimes <- sort(alltimes)
  alltimes <- alltimes - pulsetime
  alltimes <- alltimes*1000
  #diff(alltimes)
  
  gtimes[,f+1] <- alltimes
}

# get average times for each event, across subjects
avtimes = rowMeans(gtimes,na.rm=T)

eventcat <- c("Cue","Block","End","Cue","Block","Fix",
              "Cue","Block","End","Cue","Block","Fix",
              "Cue","Block","End","Cue","Block","Fix",
              "Cue","Block","End","Cue","Block","Fix",
              "Cue","Block","End","Cue","Block","Fix",
              "Cue","Block","End","Cue","Block","Fix",
              "Cue","Block","End","Cue","Block","Fix",
              "Cue","Block","End","Cue","Block","Fix")

eventnames1 <- c("Cue1","2BackFace","End1","Cue2","0BackFace","Fix1",
              "Cue3","2BackPlace","End2","Cue4","0BackFace","Fix2",
              "Cue5","0BackPlace","End3","Cue6","2BackFace","Fix3",
              "Cue7","0BackPlace","End4","Cue8","2BackPlace","Fix4",
              "Cue9","2BackPlace","End5","Cue10","0BackFace","Fix5",
              "Cue11","2BackFace","End6","Cue12","0BackPlace","Fix6",
              "Cue13","2BackFace","End7","Cue14","2BackPlace","Fix7",
              "Cue15","0BackFace","End8","Cue16","0BackPlace","Fix8")

blocknum <- c("Cue1","Block1","End1","Cue2","Block2","Fix1",
              "Cue3","Block3","End2","Cue4","Block4","Fix2",
              "Cue5","Block5","End3","Cue6","Block6","Fix3",
              "Cue7","Block7","End4","Cue8","Block8","Fix4",
              "Cue9","Block9","End5","Cue10","Block10","Fix5",
              "Cue11","Block11","End6","Cue12","Block12","Fix6",
              "Cue13","Block13","End7","Cue14","Block14","Fix7",
              "Cue15","Block15","End8","Cue16","Block16","Fix8")

# 0backface = b2, b4, b10, b15
# 0backplace = b5, b7, b12, b16
# 2backface = b1, b6, b11, b13
# 2backplace = b3, b8, b9, b14

#let's say there's a row of volume collection times d$voltimes:
#For cumulative time at each volume, compare to event times and categorize:
#Note that voltimes is time to finish, not start, each volume. This means 
#that volumes are only considered to belong to an event if they finish before the next event starts.

event=2
voleventname1 = c()
voleventname2 = c()
voleventcat = c()
volblocknum = c()

#add events to corresponding volumes
for (vol in 1:length(voltimes)) {
  if (voltimes[vol] < avtimes[event]) {
    voleventname1[vol] = eventnames1[event-1]
    voleventcat[vol] = eventcat[event-1]
    volblocknum[vol] = blocknum[event-1]
  } else {
    event = event + 1
    voleventname1[vol] = eventnames1[event-1]
    voleventcat[vol] = eventcat[event-1]
    volblocknum[vol] = blocknum[event-1]
    if (event == 49) {
      voleventname1[vol:length(voltimes)] = eventnames1[48]
      voleventcat[vol:length(voltimes)] = eventcat[48]
      volblocknum[vol:length(voltimes)] = blocknum[48]
      break
    }
  }
}

#add event labels
n$voleventname1 = voleventname1
n$voleventcat = voleventcat
n$volblocknum = volblocknum


#add stim type & wm load labels
n$stimtype = ifelse(grepl("Face",n$voleventname1),"Nface",
                    ifelse(grepl("Place",n$voleventname1),"Place",NA))

n$wmload = ifelse(grepl("0Back",n$voleventname1),"0back",                
                    ifelse(grepl("2Back",n$voleventname1),"2back",NA))

```

Write task events file
```{r}
#this contains the events for the neutral task at each volume:
write.csv(n,file = paste0("CAP_events_neut_", Sys.Date(),".csv"),row.names=FALSE)

```
