---
title: "CAP distances"
author: "Elena Peterson"
date: '2023-11-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Script for smoothing k-means distance metrics, relabelling states based on smoothed sequences, and saving brain state sequences in long (CAP_k8_long_sequence_[date].csv) and wide formats (CAP_k8_wide_smoothed_[date].csv).

Load packages & data files
```{r}
library(zoo)
library(ggplot2)

dlong = read.csv("../Data/Brain_Data/CAP_k8_long_sequence.csv",header=TRUE)
names(dlong)

dsubs = read.csv("../Data/Brain_Data/vols_per_sub.csv",header=TRUE)
names(dsubs)
```
Add volume numbers and subject labels to long data. (Each person has different numbers of starting volumes trimmed.)
```{r}

# check that lengths match
length(dlong$d1) == sum(dsubs$nvols)

# add total index
dlong$index = 1:length(dlong$d1)

# add subject labels
dlong$subid = rep(dsubs$subid, dsubs$nvols)

# add volume number per subject, and drop subject-specific number of start volumes
volnums = c()
for (num in dsubs$nvols) {
  startnum = 720 - num + 1
  volnums = c(volnums,startnum:720)
}

# lengths should match
length(volnums) == length(dlong$d1)

dlong$volnums = volnums

```

Function for computing rolling average for each person.
```{r}
# alignment is centered (for volumes 1, 2, 3: 2 is averaged with 1 and 3)

compute_roll3 = function(distcol) {
  roll_list = c()
  for (sub in dsubs$subid) {
    dsmall = dlong[dlong["subid"]==sub,]  # select rows for 1 subject
    roll3 = rollmean(dsmall[distcol],3,fill = NA)
    roll_list = c(roll_list,roll3)
  }
  return(roll_list)
}
```

Add rolling averages to long data:
```{r}
dlong$d1_roll3 = compute_roll3("d1")
dlong$d2_roll3 = compute_roll3("d2")
dlong$d3_roll3 = compute_roll3("d3")
dlong$d4_roll3 = compute_roll3("d4")
dlong$d5_roll3 = compute_roll3("d5")
dlong$d6_roll3 = compute_roll3("d6")
dlong$d7_roll3 = compute_roll3("d7")
dlong$d8_roll3 = compute_roll3("d8")

```

Reassign states based on rolling averages of distance metric:
```{r}
#subset data
all_rolls = dlong[,c("d1_roll3","d2_roll3","d3_roll3","d4_roll3","d5_roll3","d6_roll3","d7_roll3","d8_roll3")]

# this will add back in NAs
xlist = c()
for (i in 1:length(all_rolls$d1_roll3)) {
  x = which.min(all_rolls[i,]) #minimum distance is new state assignment
  x = ifelse(length(x)==1,x,NA)
  xlist = c(xlist,x)
}

# add new assignments to long data
dlong$all_states_path2 = xlist

sum(dlong$all_states_path2 == dlong$idx_raw,na.rm=T) #98091

98091/113165 # preserves 87% of the sequence

```

Visualize rolling average for each person
```{r}

# change subject as needed...
d1sub = dlong[dlong$subid=="sub-001",]

gr_sub <- ggplot(d1sub, aes(x=volnums)) + 
  geom_line(aes(y=d4_roll3),color="blue") +
  #xlab("Cylinders") + # x-axis label
  #ylab("Horsepower") + # y-axis-label
  #ggtitle("Title") + # Title of plot
  theme_classic() 

gr_sub

```

Make data frame with length = number of volumes, for visualizing state sequences.
```{r}

# non-smoothed data:
dwide_720 = data.frame("volnum"=1:720)
for (sub in dsubs$subid) {
    # get volumes belonging to sub:
    dsmall = dlong[dlong$subid==sub,c("d1","d2","d3","d4","d5","d6","d7","d8", "idx_raw")]
    names(dsmall) = paste(c("d1","d2","d3","d4","d5","d6","d7","d8", "idx_raw"), sub, sep="_")
    nvol = nrow(dsmall)
    extrarows = 720 - nvol
    dsmall[nvol+1 : extrarows, ] = NA 
    dsmall2 = rbind(dsmall[nvol+1:extrarows, ], dsmall[1:nvol,]) # put NAs at beginning
    dwide_720 = cbind(dwide_720,dsmall2)
}

# smoothed data:
dwide_720_roll3 = data.frame("volnum"=1:720)
for (sub in dsubs$subid) {
    # get volumes belonging to sub:
    dsmall = dlong[dlong$subid==sub,c("d1_roll3","d2_roll3","d3_roll3","d4_roll3","d5_roll3","d6_roll3","d7_roll3","d8_roll3", "all_states_path2")]
    names(dsmall) = paste(c("d1_roll3","d2_roll3","d3_roll3","d4_roll3","d5_roll3","d6_roll3","d7_roll3", "d8_roll3", "all_states_path2"), sub, sep="_")
    nvol = nrow(dsmall)
    extrarows = 720 - nvol
    dsmall[nvol+1:extrarows, ] = NA
    dsmall2 = rbind(dsmall[nvol+1:extrarows, ], dsmall[1:nvol,]) # put NAs at beginning
    dwide_720_roll3 = cbind(dwide_720_roll3,dsmall2)
}


# can only have numeric columns in matrices:
all_paths = dwide_720[,grep("idx_raw",names(dwide_720),value=TRUE)]
all_paths_roll3 = dwide_720_roll3[,grep("all_states_path2",names(dwide_720_roll3),value=TRUE)]

# make matrices
paths_mat = data.matrix(all_paths)
paths_r3_mat = data.matrix(all_paths_roll3)

# visualize
image(paths_mat,col=hcl.colors(8,"RdYlBu"))
image(paths_r3_mat,col=hcl.colors(8,"RdYlBu"))

# print color scheme
palette1 = data.matrix(c(1:8))
image(palette1,col=hcl.colors(8,"RdYlBu"))
```

Write files:
```{r}

write.csv(dlong, file = paste0("../Data/Brain_Data/CAP_k8_long_sequence_", Sys.Date(),".csv"),row.names=FALSE)

write.csv(dwide_720_roll3, file = paste0("../Data/Brain_Data/CAP_k8_wide_smoothed_", Sys.Date(),".csv"),row.names=FALSE)

```

