---
title: "CAP Metrics"
author: "Elena Peterson"
date: "9/2/2024"
output: html_document
---

Script for computing CAP measures from sequences of brain states

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages 
```{r}
# load packages
library(dplyr)
library(stringr)
```

Load data files 
```{r}
# PathToData = ../../Data/ScanTask_Data/

# smoothed CAP sequences, wide format
dsmooth <- read.csv("../../Data/ScanTask_Data/CAP_long_smoothed_2024-09-03.csv",header=TRUE,stringsAsFactors = FALSE)

# smoothed CAP sequences, long format
dlong <- read.csv("../../Data/ScanTask_Data/CAP_long_smoothed_2024-09-03.csv",header=TRUE,stringsAsFactors = FALSE)

# dsubs contains number of fMRI volumes per subject 
dsubs = read.csv("../../Data/ScanTask_Data/vols_per_sub.csv",header=TRUE)

# behavioral task data, contains trial-by-trial error & RT data
dtask <- read.csv("../../Data/ScanTask_Data/nback_raw_merge_2024-09-03.csv",header=TRUE)

```


```{r}

# index_to_dist: given index of events, returns sequence marking distance from events, e.g.:
# for the sequence: 0 0 0 1 0 0 1 0 0 0 0 1
# the index of events is [4, 7, 12]
# the function returns: NA NA NA 0 1 2 0 1 2 3 4 0

index_to_dist = function(vol_list) {
  nvols = 720
  allcount = rep(NA, times=vol_list[1]-1) # start with NAs before 1st event
  vol_list = c(vol_list, nvols + 1) # end with nvols plus 1
  for (i in 1:(length(vol_list)-1)) {
    start = vol_list[i] # iterate over adjacent pairs in list
    stop = vol_list[i+1]
    count1 = 0:(stop-start-1)
    allcount = c(allcount, count1)
  }
  return(allcount)
}

# index_to_dist_inv: returns inverse sequence, e.g. for the example above,
# this function returns: 3 2 1 0 2 1 0 4 3 2 1 0

index_to_dist_inv = function(vol_list) {
  nvols = 720
  allcount = c()
  vol_list = c(0, vol_list) # add zero to beginning
  for (i in 1:(length(vol_list)-1)) {
    start = vol_list[i] # iterate over adjacent pairs in list
    stop = vol_list[i+1]
    count1 = (stop-start-1):0
    allcount = c(allcount, count1)
  }
  # add NAs after last event
  na_num = nvols - vol_list[length(vol_list)]
  na_end =rep(NA, times=na_num)
  allcount = c(allcount, na_end)
  return(allcount)
}
```


Use long data to track behavioral lapses & state transitions
```{r}
# make sure dlong rows in correct order:
dlong <- dlong[order(dlong$subid, dlong$volnums),]

# intialize error & long RT variables
dlong$errors = 0
dlong$long_rts = 0

# track errors & long RTs for each sub (including non-responses)
for (sub in dsubs$subid) {
  
  # look at subject's data
  dsmall = dtask[dtask$subid == sub,]
  
  # pulsetime is task starting time
  pulsetime <- dsmall$trial_startT[3]
  
  # compute times errors occur, convert to volumes, add to dlong
  error_times = dsmall$trial_startT[dsmall$key_resp_3.corr==0] - pulsetime
  error_vols = floor(error_times/.8)
  dlong$errors[dlong$subid==sub & dlong$volnum %in% error_vols] = 1
  
  # get subject mean and sd to identify long RTs
  sub_sd = sd(dsmall$key_resp_3.rt, na.rm=T)
  sub_mean = mean(dsmall$key_resp_3.rt, na.rm=T)
  
  # get RTs > 2 sd above mean, compute timing and convert to volume index, add to dlong
  long_rts = dsmall$key_resp_3.rt[which(dsmall$key_resp_3.rt > (sub_mean+2*sub_sd))]
  trial_times = dsmall$trial_startT[which(dsmall$key_resp_3.rt > (sub_mean+2*sub_sd))]
  rt_times = long_rts + trial_times - pulsetime
  rt_vols = floor(rt_times/.8)
  dlong$long_rts[dlong$subid==sub & dlong$volnum %in% rt_vols] = 1
  
}

# convert binary indices to continous metrics

for (sub in dsubs$subid) {
  # look at subject's data
  dsmall = dlong[dlong$subid == sub,]
  vol1 = dsmall$volnums[1] # starting vol number
  errorlist = dsmall$volnums[which(dsmall$errors == 1)]
  
  if (length(errorlist) > 0) {
    list720 = index_to_dist(errorlist)
    dlong$errors_dist[dlong$subid==sub] = list720[vol1:720]
    
    # do inverse as well
    list720inv = index_to_dist_inv(errorlist)
    dlong$errors_dist_inv[dlong$subid==sub] = list720inv[vol1:720]
    
  } else if (length(errorlist) == 0) {
      next
  }
}


for (sub in dsubs$subid) {
  # look at subject's data
  dsmall = dlong[dlong$subid == sub,]
  vol1 = dsmall$volnums[1]
  rtlist = dsmall$volnums[which(dsmall$long_rts == 1)]
  
  if (length(rtlist) > 0) {
      list720 = index_to_dist(rtlist)
      dlong$long_rts_dist[dlong$subid==sub] = list720[vol1:720]
      
      # do inverse as well
      list720inv = index_to_dist_inv(rtlist)
      dlong$long_rts_dist_inv[dlong$subid==sub] = list720inv[vol1:720]
     
  } else if (length(rtlist) == 0) {
      next
  } 
}

```

Merge dsmooth event labels with dlong
```{r}
dsmooth$volnums = dsmooth$volnum
dsmooth_events = subset(dsmooth, select = c(voltimes, voleventname1, voleventcat, volblocknum, stimtype, wmload, volnums))

dlong2 = merge(dlong,dsmooth_events,sort=FALSE)
dlong2 = arrange(dlong2,subid)

```

```{r}
#this file contains the volume by volume events in long form...
write.csv(dlong2,file = paste0("../../Data/ScanTask_Data/CAP_k8_long_metrics_", Sys.Date(),".csv"),row.names=FALSE)
```


Create new wide data frame
```{r}
# new df for saving metrics by sub
dwide = data.frame("subid"=dsubs$subid)

# or previous dataset
#dwide = read.csv("../../Data/Merged_Data/Merged_Data_Wide_2024-04-16.csv",stringsAsFactors = FALSE, header = TRUE)
dim(dwide)
names(dwide)
```

Function for computing time-in-state by condition (proportional)
```{r}
# define function for computing time-in-state by condition (proportional)

compute_ts1 = function(event, state) {
  tslist = c()
  for (sub in dsubs$subid) {
    # subset data by subject & task event type
    df2 = dlong2[dlong2["subid"]==sub & 
                grepl(event,dlong2$voleventname1) & 
                !is.na(dlong2["all_states_path2"]),]  # select rows for 1 subject
    nvol = length(df2$subid)
    blist = which(df2$all_states_path2==state)
    # outcome = number of state volumes out of total event volumes:
    ts = length(blist)/nvol
    tslist = c(tslist,ts)
  }
  return(tslist)
}
```

Function for computing number of transitions by condition
```{r}
# define function for computing unidirectional state 1 to state 2 transitions by condition

compute_trx = function(event, state1, state2) {
  trxlist = c()
  for (sub in dsubs$subid) {
    # subset data by subject & task event type
    df2 = dlong2[dlong2["subid"]==sub & 
                grepl(event,dlong2$voleventname1) & 
                !is.na(dlong2["all_states_path2"]),]  # select rows for 1 subject
    # get indices of state 1 (+1) and state 2, then count overlap
    state1_idx = df2$volnum[which(df2$all_states_path2==state1)] + 1
    state2_idx = df2$volnum[which(df2$all_states_path2==state2)]
    trxcount = sum(state1_idx %in% state2_idx)
    trxlist = c(trxlist,trxcount)
    }  
  return(trxlist)
}

```


Function for computing persistence by block (not "true" persistence because we're chopping up sequence)
```{r}

# look at persistence within a specific window...
# use this when you want to look at persistence within non-adjacent blocks, I think...
compute_ps1_block = function(event,state) {
  
  pslist = c() # initialize output list
  eventlist = unique(dsmooth$volblocknum[grep(event, dsmooth$voleventname1)]) # get block IDs for event type
  
  for (sub in dsubs$subid) {
    
    xlist = c() # initialize subject list of persistences
    
    # subset data by subject
    df1 = dlong2[dlong2["subid"]==sub & 
                !is.na(dlong2["all_states_path2"]),]
    
    # if state is absent entirely move on to next person:
    if (!(state %in% df1$all_states_path2)) {
      ps = NA
      next
    }
    
    # iterate through block IDs:
    for (ev in eventlist) {
      
      # subset again by block number
      df2 = df1[df1["volblocknum"]==ev,]
    
      #if brain state is absent entirely, then move on to next block
      if (sum(df2$all_states_path2==state)==0) {
        next
      }
      
      # compute persistence:
      x = 0 # initialize vol count
      for (vol in 1:nrow(df2)) { # for each volume
        if (df2$all_states_path2[vol]==state) { # count if it's the brain state of interest
          x = x+1
          if (vol==nrow(df2)) { # if it's the last volume, make sure to record it
            xlist = c(xlist,x)
          }
        } else if (x!=0) { # if the brain state changes, record the duration and move on
          xlist = c(xlist,x)
          x=0
        }
      }
    }
    
    # if a state is absent from all blocks, record as NA
    if (length(xlist)==0) {
      ps = NA
    } else {
      #get average persistence for all blocks for one sub:
      ps = mean(xlist)
    }
    pslist = c(pslist,ps)
  }
  
  return(pslist)
}

```

Compute persistence overall
```{r}

#Compute persistence overall

compute_ps1_overall = function(state) {
  pslist = c()
  for (sub in dsubs$subid) { #for each subject
    
    xlist = c() 
    
    # subset data by subject
    df2 = dlong2[dlong2["subid"]==sub & 
                !is.na(dlong2["all_states_path2"]),]
    
    # if state is absent entirely move on to next person:
    if (!(state %in% df2$all_states_path2)) {
      ps = NA
      pslist = c(pslist,ps)
      next
    }
    
    # compute persistence:
    x = 0
    for (vol in 1:nrow(df2)) { 
      if (df2$all_states_path2[vol]==state) { # count if it's the brain state of interest
        x = x+1
        if (vol==nrow(df2)) { # if it's the last volume, make sure to record it
            xlist = c(xlist,x)
        }
      } else if (x!=0) { # if the brain state changes, record the duration and move on
          xlist = c(xlist,x)
          x=0
        }
    }
   ps = mean(xlist)
   pslist = c(pslist,ps)
}
  return(pslist)
}

```

Add overall CAP measures to wide format file.
Takes a few minutes.
```{r}

for (bs in 1:8) {
  print(paste("Computing measures for State", bs))
  # compute time-in-state measures:
  dwide[[paste0("overall_ts_", bs)]] = compute_ts1("*", state = bs)
  dwide[[paste0("ts_", bs, "_0back")]] = compute_ts1("0Back", state = bs)
  dwide[[paste0("ts_", bs, "_2back")]] = compute_ts1("2Back", state = bs)
  dwide[[paste0("ts_", bs, "_fix")]] = compute_ts1("Fix", state = bs)
  
  # compute persistence measures:
  dwide[[paste0("overall_ps_", bs)]] = compute_ps1_overall(state = bs) # issues with state 8
  dwide[[paste0("ps_", bs, "_0back")]] = compute_ps1_block("0Back", state = bs)
  dwide[[paste0("ps_", bs, "_2back")]] = compute_ps1_block("2Back", state = bs)
  dwide[[paste0("ps_", bs, "_fix")]] = compute_ps1_block("Fix", state = bs)
}

# compute transition measures (just interested in state 6 to state 1 transitions for now)
dwide$trx_61 = compute_trx("*",state1 = 6, state2 = 1)
dwide$trx_61_0back = compute_trx("0Back",state1 = 6, state2 = 1)
dwide$trx_61_2back = compute_trx("0Back",state1 = 6, state2 = 1)

```

Write out files
```{r}
#this file contains the overall time-in-state, persistence, and transition measures for each person.
write.csv(dwide,file = paste0("../../Data/ScanTask_Data/CAP_measures_overall_", Sys.Date(),".csv"),row.names=FALSE)

```
