---
title: "MRI n-back"
author: "Elena Peterson"
date: "8/12/2024"
output: html_document
---

Script for extracting behavioral measures from fmri task.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages & data files 
```{r}
mydata <- read.csv("../../Data/Brain_Data/nback_raw_merge_2024-08-12.csv",header=TRUE)

# subject list
dsubs = read.csv("../../Data/Brain_Data/vols_per_sub.csv",header=TRUE)
names(dsubs)

# load packages
library(dplyr)
library(stringr)
```

Clean up mydata for computing accuracy measures
```{r}
mydata$subid = paste0("sub-", substr(mydata$participant, 5,7))

# remove 3 rows at start of task file
mydata_acc <- mydata[-which(mydata$BlockFile==""),]

#remove rows in between blocks
mydata_acc <- mydata_acc[!is.na(mydata_acc$trials.thisN),] #removes rows in between blocks
#mydata <- mydata[!is.na(mydata$key_resp_3.corr),]

# remove 1st 2 trials of 2-back blocks
mydata_acc <- mydata_acc[-which(mydata_acc$BlockType == "2-Back" & mydata_acc$trials.thisIndex <= 1),]

#remove rows with missing responses (could try both ways)
mydata_acc_NoMiss <- mydata_acc[-which(mydata_acc$key_resp_3.keys=="None"),]
```

Get overall accuracy measures
```{r}
#note: aggregate reorders IDs. So to keep it the same, first get unique IDs, then sort.

master <- data.frame("participant" = sort(unique(mydata$participant)))


master$OverallAccuracy <- aggregate(list(OverallAccuracy=mydata_acc$key_resp_3.corr),
                                list(participant=mydata_acc$participant),
                                mean)$OverallAccuracy

master$OverallNoAccuracy <- aggregate(list(OverallNoAccuracy=mydata_acc[mydata_acc$TargetType == "nonlure",]$key_resp_3.corr),list(participant=mydata_acc[mydata_acc$TargetType == "nonlure",]$participant),mean)$OverallNoAccuracy

master$OverallYesAccuracy <- aggregate(list(OverallYesAccuracy=mydata_acc[mydata_acc$TargetType == "target",]$key_resp_3.corr),list(participant=mydata_acc[mydata_acc$TargetType == "target",]$participant), mean)$OverallYesAccuracy

master$OverallLureAccuracy <- aggregate(list(OverallLureAccuracy=mydata_acc[mydata_acc$TargetType == "lure",]$key_resp_3.corr),list(participant=mydata_acc[mydata_acc$TargetType == "lure",]$participant), mean)$OverallLureAccuracy

master$Zeroback_Accuracy <- aggregate(list(Zeroback_Accuracy=mydata_acc[mydata_acc$BlockType == "0-Back",]$key_resp_3.corr),
                                list(participant=mydata_acc[mydata_acc$BlockType == "0-Back",]$participant),
                                mean)$Zeroback_Accuracy

master$Twoback_Accuracy <- aggregate(list(Twoback_Accuracy=mydata_acc[mydata_acc$BlockType == "2-Back",]$key_resp_3.corr),
                                list(participant=mydata_acc[mydata_acc$BlockType == "2-Back",]$participant),
                                mean)$Twoback_Accuracy

# check who has low accuracy scores
master$Participant[which(master$OverallAccuracy<.6)]

# check values of low accuracy
#master$OverallAccuracy[which(master$OverallAccuracy<.6)] 

```


Get overall RT measures
```{r}
#note: aggregate reorders IDs. So to keep it the same, first trim IDs to be consistent,
# then sort.

master$OverallRT <- aggregate(list(OverallRT=mydata_acc$key_resp_3.rt),
                                list(participant=mydata_acc$participant),
                                mean,na.rm=T)$OverallRT

master$Zeroback_RT <- aggregate(list(ZerobackRT=mydata_acc[mydata_acc$BlockType == "0-Back",]$key_resp_3.rt),
                                list(participant=mydata_acc[mydata_acc$BlockType == "0-Back",]$participant),
                                mean,na.rm=T)$ZerobackRT

master$Twoback_RT <- aggregate(list(TwobackRT=mydata_acc[mydata_acc$BlockType == "2-Back",]$key_resp_3.rt),
                                list(participant=mydata_acc[mydata_acc$BlockType == "2-Back",]$participant),
                                mean,na.rm=T)$TwobackRT

master$Overall_logRT = log(master$OverallRT*1000)
master$Zeroback_logRT = log(master$Zeroback_RT*1000)
master$Twoback_logRT = log(master$Twoback_RT*1000)

```


Define function for getting different accuracy types by condition
```{r}

#count missing responses
missing <- function(y) {
  sum(y=="None")
}

#accuracy by condition function
which_acc <- function(condition) {
  
  Acc <- aggregate(list(BlockAccuracy=mydata_acc[condition,]$key_resp_3.corr),
  list(participant=mydata_acc[condition,]$participant),mean)
  
  Miss <- aggregate(list(Misses=mydata_acc[condition,]$key_resp_3.keys),
  list(participant=mydata_acc[condition,]$participant),missing)

  merge(Acc,Miss,all=T)
}

```

Check for lots of missing responses (at least half per block)
```{r}
nblocks = paste0("Blocks/","Block",1:16,".xlsx")

# get block level accuracy & missed responses
for (b in nblocks) {
  B1 <- which_acc(mydata_acc$BlockFile==b)
  if (sum(B1$Misses>4)>0) {
    print(B1$participant[which(B1$Misses>4)])
  }
}

```

Add more specific accuracy measures, starting with 0-back condition:
```{r}
master$Zeroback_Misses <- aggregate(
  list(Zeroback_Misses=mydata_acc[mydata_acc$BlockType == "0-Back",]$key_resp_3.keys), list(participant=mydata_acc[mydata_acc$BlockType == "0-Back",]$participant), missing)$Zeroback_Misses

master$Zeroback_Lures <- aggregate(
  list(Zeroback_Lures=mydata_acc[mydata_acc$BlockType == "0-Back" & mydata_acc$TargetType=="lure",]$key_resp_3.corr), list(participant=mydata_acc[mydata_acc$BlockType == "0-Back"& mydata_acc$TargetType=="lure",]$participant), mean)$Zeroback_Lures

master$Zeroback_Lures <- aggregate(
  list(Zeroback_Lures=mydata_acc[mydata_acc$BlockType == "0-Back" & mydata_acc$TargetType=="lure",]$key_resp_3.corr), list(participant=mydata_acc[mydata_acc$BlockType == "0-Back"& mydata_acc$TargetType=="lure",]$participant), mean)$Zeroback_Lures

master$Zeroback_Nonlures <- aggregate(
  list(Zeroback_Nonlures=mydata_acc[mydata_acc$BlockType == "0-Back" & mydata_acc$TargetType=="nonlure",]$key_resp_3.corr), list(participant=mydata_acc[mydata_acc$BlockType == "0-Back"& mydata_acc$TargetType=="nonlure",]$participant), mean)$Zeroback_Nonlures

master$Zeroback_Targets <- aggregate(
  list(Zeroback_Targets=mydata_acc[mydata_acc$BlockType == "0-Back" & mydata_acc$TargetType=="target",]$key_resp_3.corr), list(participant=mydata_acc[mydata_acc$BlockType == "0-Back"& mydata_acc$TargetType=="target",]$participant), mean)$Zeroback_Targets
```

Save behavioral measures (acc & RT)
```{r}
write.csv(master, paste0("../../Data/Brain_Data/Task_Bx_",Sys.Date(),".csv"),row.names = FALSE)
```

