---
title: "MRI n-back"
author: "Elena Peterson"
date: "8/12/2024"
output: html_document
---

Script for extracting behavioral measures from fmri task.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages & data files 
```{r}
mydata <- read.csv("../../Data/ScanTask_Data/nback_raw_merge_2024-09-03.csv",header=TRUE)

# subject list
dsubs = read.csv("../../Data/ScanTask_Data/vols_per_sub.csv",header=TRUE)
names(dsubs)

# load packages
library(dplyr)
library(stringr)
```

Clean up mydata for computing accuracy measures
```{r}
mydata$subid = paste0("sub-", substr(mydata$participant, 5,7))

# remove 3 rows at start of task file
mydata_acc <- mydata[-which(mydata$BlockFile==""),]

#remove rows in between blocks
mydata_acc <- mydata_acc[!is.na(mydata_acc$trials.thisN),] #removes rows in between blocks
#mydata <- mydata[!is.na(mydata$key_resp_3.corr),]

# remove 1st 2 trials of 2-back blocks
mydata_acc <- mydata_acc[-which(mydata_acc$BlockType == "2-Back" & mydata_acc$trials.thisIndex <= 1),]

#remove rows with missing responses (could try both ways)
mydata_acc_NoMiss <- mydata_acc[-which(mydata_acc$key_resp_3.keys=="None"),]
```

Get overall accuracy measures
```{r}

master <- data.frame("participant" = sort(unique(mydata$participant)))

# Calculate Overall Accuracy
master <- mydata_acc %>%
  group_by(participant) %>%
  summarise(OverallAccuracy = mean(key_resp_3.corr)) %>%
  right_join(master, by = "participant")

# Calculate Overall No Accuracy
master <- mydata_acc %>%
  filter(TargetType == "nonlure") %>%
  group_by(participant) %>%
  summarise(OverallNoAccuracy = mean(key_resp_3.corr)) %>%
  right_join(master, by = "participant")

# Calculate Overall Yes Accuracy
master <- mydata_acc %>%
  filter(TargetType == "target") %>%
  group_by(participant) %>%
  summarise(OverallYesAccuracy = mean(key_resp_3.corr)) %>%
  right_join(master, by = "participant")

# Calculate Overall Lure Accuracy
master <- mydata_acc %>%
  filter(TargetType == "lure") %>%
  group_by(participant) %>%
  summarise(OverallLureAccuracy = mean(key_resp_3.corr)) %>%
  right_join(master, by = "participant")

# Calculate Overall Zero-back Accuracy
master <- mydata_acc %>%
  filter(BlockType == "0-Back") %>%
  group_by(participant) %>%
  summarise(ZeroBack_Accuracy = mean(key_resp_3.corr)) %>%
  right_join(master, by = "participant")

# Calculate Overall Two-back Accuracy
master <- mydata_acc %>%
  filter(BlockType == "2-Back") %>%
  group_by(participant) %>%
  summarise(TwoBack_Accuracy = mean(key_resp_3.corr)) %>%
  right_join(master, by = "participant")

```

```{r}
# check who has low accuracy scores
master$participant[which(master$OverallAccuracy<.6)]

# check values of low accuracy
master$OverallAccuracy[which(master$OverallAccuracy<.6)] 

```


Get overall RT measures
```{r}
#note: aggregate reorders IDs. So to keep it the same, first trim IDs to be consistent,
# then sort.

# Calculate Overall RT
master <- mydata_acc %>%
  group_by(participant) %>%
  summarise(OverallRT = mean(key_resp_3.rt, na.rm=T)) %>%
  right_join(master, by = "participant")

# Calculate Zero-back RT
master <- mydata_acc %>%
  filter(BlockType == "0-Back") %>%
  group_by(participant) %>%
  summarise(Zeroback_RT = mean(key_resp_3.rt, na.rm=T)) %>%
  right_join(master, by = "participant")

# Calculate Two-back RT
master <- mydata_acc %>%
  filter(BlockType == "2-Back") %>%
  group_by(participant) %>%
  summarise(Twoback_RT = mean(key_resp_3.rt, na.rm=T)) %>%
  right_join(master, by = "participant")

# convert to ms & log transform
master$Overall_logRT = log(master$OverallRT*1000)
master$Zeroback_logRT = log(master$Zeroback_RT*1000)
master$Twoback_logRT = log(master$Twoback_RT*1000)

```


Define function for getting different accuracy types by condition
```{r}

#count missing responses
missing <- function(y) {
  sum(y=="None")
}

#accuracy by condition function
which_acc <- function(condition) {
  
  Acc <- aggregate(list(BlockAccuracy=mydata_acc[condition,]$key_resp_3.corr),
                   list(participant=mydata_acc[condition,]$participant),mean)
  
  Miss <- aggregate(list(Misses=mydata_acc[condition,]$key_resp_3.keys),
                    list(participant=mydata_acc[condition,]$participant),missing)

  merge(Acc,Miss,all=T)
}

```

Check for lots of missing responses (at least half per block)
```{r}
nblocks = paste0("Blocks/","Block",1:16,".xlsx")

# get block level accuracy & missed responses
for (b in nblocks) {
  B1 <- which_acc(mydata_acc$BlockFile==b)
  if (sum(B1$Misses>4)>0) {
    print(B1$participant[which(B1$Misses>4)])
  }
}

```

Add more specific accuracy measures, starting with 0-back condition:
```{r}
# Count Zero-back Misses
master <- mydata_acc %>%
  filter(BlockType == "0-Back") %>%
  group_by(participant) %>%
  summarise(Zeroback_Misses = missing(key_resp_3.keys)) %>%
  right_join(master, by = "participant")

# Zero-back Lure Accuracy
master <- mydata_acc %>%
  filter(BlockType == "0-Back") %>%
  filter(TargetType == "lure") %>%
  group_by(participant) %>%
  summarise(Zeroback_Lures = mean(key_resp_3.corr)) %>%
  right_join(master, by = "participant")

# Zero-back Non-Lure Accuracy
master <- mydata_acc %>%
  filter(BlockType == "0-Back") %>%
  filter(TargetType == "nonlure") %>%
  group_by(participant) %>%
  summarise(Zeroback_Nonlures = mean(key_resp_3.corr)) %>%
  right_join(master, by = "participant")

# Zero-back Target Accuracy
master <- mydata_acc %>%
  filter(BlockType == "0-Back") %>%
  filter(TargetType == "target") %>%
  group_by(participant) %>%
  summarise(Zeroback_Targets = mean(key_resp_3.corr)) %>%
  right_join(master, by = "participant")


```

Save behavioral measures (acc & RT)
```{r}
write.csv(master, paste0("../../Data/ScanTask_Data/Task_Bx_",Sys.Date(),".csv"),row.names = FALSE)
```

